Efficiently Handling Guards in the Low Level Design of RPython's tracing JIT

### 2012-08-06 talk/vmil2012/paper.tex
~add a pie chart to the appendix showing the distribution of operations before and after optimization (hmmmm pie)
refactor another evaluation paragraph and mark pending tasks for the evaluation section

The overhead that is incurred by the JIT to manage the \texttt{resume data},
the \texttt{low-level resume data} and the generated machine code is shown in
Figure~\ref{fig:backend_data}. It shows the total memory consumption of the
code and of the data generated by the machine code backend for the different
benchmarks mentioned above. The size of the machine code is composed of the
size of the compiled operations, the trampolines generated for the guards and a
set of support functions that are generated when the JIT starts and are shared
by all compiled traces. The size of the \texttt{low-level resume data} is the
size of the registers and stack to IR-level variable mappings and finally the
size of the \texttt{resume data} is an approximation of the size of the
compressed high-level resume data. While the \texttt{low-level resume data} has
a size of about 15\% to 20\% of the generated instructions the \texttt{resume
data} is even in the compressed form larger than the generated machine code.

Tracing JITs compilers only compile a subset of the executed program so the
amount of generated machine code will be smaller than for function based JITs.
At the same time there is a several times larger overhead for keeping the
resume information for the guards. The generated machine code accounts for
20.21\% to 37.97\% of the size required for storing the different kinds of
resume data.

@@

* Evaluation
   * Measure guard memory consumption and machine code size
   * Extrapolate memory consumption for guard other guard encodings
      * compare to naive variant
   * Measure how many guards survive optimization
   * Measure the of guards and how many of these ever fail

@@

\section*{Acknowledgements}
\section*{Appendix}
\begin{figure*}
    \include{figures/ops_count_table}
    \caption{Relative numbers of operations in the traces generated for
    different benchmarks}
    \label{fig:ops_count}
\end{figure*}
\bibliographystyle{abbrv}
\bibliography{zotero,paper}
\listoftodos
\end{document}

>>>>

The overhead that is incurred by the JIT to manage the \texttt{resume data},
the \texttt{low-level resume data} as well as the generated machine code is
shown in Figure~\ref{fig:backend_data}. It shows the total memory consumption
of the code and of the data generated by the machine code backend and an
approximation of the size of the \texttt{resume data} structures for the
different benchmarks mentioned above. The size of the machine code is composed
of the size of the compiled operations, the trampolines generated for the
guards and a set of support functions that are generated when the JIT starts
and are shared by all compiled traces. The size of the \texttt{low-level resume
data} is the size of the compressed mapping from registers and stack to
IR-level variable and finally the size of the \texttt{resume data} is an
approximation of the size of the compressed high-level resume data\todo{explain
why it is an approximation}.

Compared to the size of the generated machine code the compressed
\texttt{low-level resume data} is about 15\% to 20\% of that size, depending on
the benchmark. On the other hand the generated machine code has only a size
ranging from 20.21\% to 37.98\% of the size of the high and low-level
\texttt{resume data} being compressed as described before.

Tracing JIT compilers only compile the subset of the code executed in a program
that is traced in a hot loop, for this reason the amount of generated machine
code will be smaller than in other juts-in-time compilation approaches. Still
the overhead associated to guards to resume execution from a side exit appears
to be high.\bivab{put into relation to other JITs, compilers in general}

@@

\todo{compare to naive variant of resume data}
\todo{Measure the of guards and how many of these ever fail}

@@

\section*{Acknowledgements}
\section*{Appendix}
\begin{figure*}
    \include{figures/ops_count_table}
    \caption{Relative numbers of operations in the traces generated for
    different benchmarks}
    \label{fig:ops_count}
\end{figure*}
\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/ops_pie.pdf}
\caption{Relative frequency of operations before and after optimization}
\label{fig:ops_pie}
\end{figure*}
\bibliographystyle{abbrv}
\bibliography{zotero,paper}
\listoftodos
\end{document}
